---
title: "Sustainability Analysis in Trail Running Events"
author: "Cristi M."
date: "2025-06"
format: html
editor: visual
---

## 🧭 Working Directory in R

R needs to know *where* to look for your data files and *where* to save output. This location is your **working directory**.

- `getwd()` shows your current directory.
- `setwd("your/path")` changes it.

> ✅ Tip: Always set it at the start of your project.

```{r}
getwd()
#setwd("D:/proiecte cristi/Prezentari Internationale/SES TRAIL/data for trail running")
```

---

## 📦 Packages for Trail Running Sustainability Analysis

We use different file types:
- **Text** (event descriptions & regulations)
- **ZIP** (compressed datasets)
- **Spatial** (GPX, shapefiles)

### Text Packages

| Package      | Use                                                  |
|--------------|------------------------------------------------------|
| `readtext`   | Read `.txt`, `.docx`, `.pdf` into data frames        |
| `tm`         | Text mining & preprocessing                          |
| `SnowballC`  | Word stemming                                        |

### ZIP Utilities

- `unzip()` → extract zip content
- `list.files()` → list all files inside folders

### Spatial Packages

| Package   | Purpose                                                  |
|-----------|----------------------------------------------------------|
| `sf`      | Handle spatial vector data                               |
| `tmap`    | Thematic mapping                                         |
| `leaflet` | Interactive maps                                         |
| `raster`  | Work with elevation/imagery (optional)                   |

```{r}
# install.packages(c("readtext", "tidytext", "stopwords", "wordcloud", "RColorBrewer"))
```

---

## 📚 What Is an R Package?

An **R package** is a toolbox of functions, data, and documentation.  
Think of them as extensions to R.

### Example

```{r}
# install.packages("sf")    
library(sf)
```

---

## 🔓 Load and Process Text Data from ZIP

```{r}
# Step 1: Select zip file interactively
zip_path <- file.choose()

# Step 2: Unzip to folder (preserve subfolders)
unzip(zip_path, exdir = "unzipped_data")

# Step 3: List all .txt and .docx files recursively
text_files <- list.files("unzipped_data", pattern = regex("\\.(txt|docx)$"),
                         recursive = TRUE, full.names = TRUE)

# Step 4: Keep only valid files
text_files <- text_files[file.exists(text_files)]

# Step 5: Load texts
library(readtext)
text_data <- readtext(text_files)

# Step 6: Preview
head(text_data)
```

---

## 🧹 Text Cleaning and Tokenization

```{r}
library(dplyr)
library(tidytext)
library(stringr)
library(stopwords)

# Step 1: Break into words
tidy_text <- text_data %>%
  select(doc_id, text) %>%
  unnest_tokens(word, text)

# Step 2: Remove Romanian stopwords, numbers, short words
romanian_stopwords <- stopwords("ro")

tidy_text <- tidy_text %>%
  filter(!word %in% romanian_stopwords) %>%
  filter(!str_detect(word, "^[0-9]+$")) %>%
  filter(str_length(word) > 2)

# Step 3: Preview
head(tidy_text)
```

---

## ☁️ Generate a Word Cloud

```{r}
library(wordcloud)
library(RColorBrewer)

# Word frequency
word_freq <- tidy_text %>%
  count(word, sort = TRUE)

# Preview top 10
head(word_freq, 10)

# Word cloud
set.seed(1234)
wordcloud(words = word_freq$word,
          freq = word_freq$n,
          min.freq = 2,
          max.words = 150,
          random.order = FALSE,
          rot.per = 0.25,
          colors = brewer.pal(8, "Dark2"))
```

---

## ✅ Summary

- Load ZIP with `.docx` or `.txt` content
- Clean, tokenize and remove stopwords
- Visualize word importance with a cloud

🎯 Next: Analyze **regulations** or **map GPX routes** using `sf`.
